Classical Mechanics Assignment 8

### 2015-11-23 classical/latex/a8.tex
Add stuffs.

\title{Classical Mechanics Assignment 6}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One: Ball on an Inclined Plane}
\textbf{Write the Lagrangian for a cylinder (mass m, radius R, and moment of inertia I) that rolls without slipping straight down an inclined plane at an angle $\alpha$ from the horizontal. Use as your generalized coordinates the cylinder’s distance x measured down the plane from its starting point. Solve the Lagrange equation for the cylinder’s acceleration $\ddot{x}$. Remember that $T = \frac{1}{2} mv^2 + \frac{1}{2} I\omega^2$ where $v$ is the velocity of the center of mass and $\omega$ is the cylinder’s angular velocity.}

\begin{align*}
  x' &= x\cos(\alpha)\\
  y' &= -x\sin(\alpha)\\
  \omega &= \frac{v}{2\pi R}\\
  \vspace{1cm}\\
  U &= mgy' = -mgx\sin(\alpha)\\
  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
  &= \frac{1}{2}m\left(\dot{x}'^2 + \dot{y}'^2\right) + \frac{1}{2}I\left(\frac{\dot{x}'^2 + \dot{y}'^2}{2\pi R}\right)\\
  &= \frac{1}{2}m\left(\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)\right) + \frac{1}{2}I\left(\frac{\dot{x}^2\sin^2(\alpha) + \dot{x}^2\cos^2(\alpha)}{4\pi^2R^2}\right)\\
  &= m\left(\dot{x}^2\right) + I\left(\frac{\dot{x}^2}{4\pi^2R^2}\right)\\
  &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right)\\
  L &= \dot{x}^2\left(\frac{I + 4\pi^2R^2m}{4\pi^2R^2}\right) + mgx\sin(\alpha)\\
  \vspace{1cm}\\
  &\frac{dL}{dx} - \frac{d}{dt}\frac{dL}{d\dot{x}} = 0\\
  &mg\sin(\alpha) - \ddot{x}\left(\frac{I + 4\pi^2R^2m}{\pi^2R^2}\right) = 0\\
  \ddot{x} &= \frac{\pi^2R^2mg\sin(\alpha)}{I + 4\pi^2R^2m}\\
  x &= \frac{\pi^2R^2mg\sin(\alpha)}{2I + 8\pi^2R^2m}t^2
\end{align*}

It took me a while to figure out why I needed to set up kinetic energy in cartesian coordinates. I thought I could just take my axes to be in the x direction (down the slope), but that isn't how we've done other problems. Although I did use the cartesian coordinates method, I found that using other cartesian coordinates (ones more natural to the problem, where x is in the direction of the wedge) would have yielded the same answer. I will try to apply these smarter cartesian coordinates to further problems.\\

\section{Problem Two: Bead on a Spiral}
\textbf{A smooth wire is bent into the shape of a helix, with cylindrical coordinates $\rho=R$ and $z=\lambda\phi$, where $R$ and $\lambda$ are constants and the z axis is vertically up (and gravity is vertically down). Using z as your generalized coordinate, write down the Lagrangian for a bead of mass m threaded on the wire. Find the Lagrange equation and the bead’s vertical acceleration $\ddot{z}$. In the limit that $R → 0$, what is $\ddot{z}$? Does this make sense?}

\begin{align*}
  v &= \sqrt{\dot{\rho}^2 + \rho^2\dot{\phi}^2 + \dot{z}^2}\\
  &= \sqrt{\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}mv^2\\
  &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right)\\
  U &= mgz\\
  L &= \frac{1}{2}m\left(\rho^2\frac{\dot{z}^2}{\lambda^2} + \dot{z}^2\right) - mgz\\
  \vspace{1cm}\\
  &\frac{dL}{dz} - \frac{d}{dt}\frac{dL}{d\dot{z}} = 0\\
  -mg - \ddot{z}m\left(\frac{\rho^2}{\lambda^2}+1\right) &= 0\\
  \ddot{z} &= \frac{-g}{\frac{R^2}{\lambda^2}+1}\\
\end{align*}

Fun problem. It was simple and easy to set up. It was interesting to think about how the problem would be different if the helix was moving. Also, when R goes to zero, the rhs of the equation just turns to -g, which is exactly what you would expect: a frictionless bead falling down a straight loop. It was cool that this works out.\\

\section{Problem Three: }
\textbf{Write the Lagrangian for a cylinder (mass m, radius r, and moment of inertia $I=\frac{mr^2}{2}$) that rolls without slipping in the bottom of a circular shaped hole of equation $z = −\sqrt{R^2 − x^2}$ , where z is the vertical axis (see figure below). Find the angular frequency of small oscillations about the bottom of the hole.}

\begin{align*}
  x &= (R-r)\sin(\theta)\\
  z &= -(R-r)\cos(\theta)\\
  v^2 &= (R-r)^2\dot{\theta}^2\\
  \omega^2 &= \frac{v^2}{r^2} = \frac{(R-r)^2\dot{\theta}^2}{r^2}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}mv^2 + \frac{1}{2}I\omega^2\\
  &= \frac{1}{2}m(R-r)^2\dot{\theta}^2 + \frac{1}{2}I\frac{(R-r)^2\dot{\theta}^2}{r^2}\\
  &= \frac{1}{2}\left(m + \frac{I}{r^2}\right)(R-r)^2\dot{\theta}^2\\
  U &= -mg(R-r)\cos(\theta)\\
  L &= \frac{1}{2}\left(m + \frac{I}{r^2}\right)(R-r)^2\dot{\theta}^2 + mg(R-r)\cos(\theta)\\
  \vspace{1cm}\\
  \frac{dL}{d\theta} - \frac{d}{dt}\frac{dL}{d\dot{\theta}} &= 0\\
  -mg(R-r)\sin(\theta) - \left(m + \frac{I}{r^2}\right)(R-r)^2\ddot{\theta} &= 0\\
  \left(m + \frac{I}{r^2}\right)(R-r)\ddot{\theta} &= -mg\sin(\theta)\\
  \ddot{\theta} &= \frac{-mg\sin(\theta)}{\left(m + \frac{I}{r^2}\right)(R-r)}\\
  \ddot{\theta} &= \frac{-2g\sin(\theta)}{3(R-r)}\\
\end{align*}

We then take the small angle approximation:

\begin{align*}
  \ddot{\theta} &= \frac{-2g\theta}{3(R-r)}\\
\end{align*}

Suppose our solution for $\theta$ takes the form $\theta = e^{wt}$. Then we can see that $\omega = \sqrt{\frac{2g}{3(R-r)}}$.\\

Cool problem! It felt nice to get a final result that looks like that of a pendulum, even though we used a completely different method. The small angle approximation helped a lot for this problem, I was struggling at applying it for a while. Remembering how to find the angular frequency was good for me, the method of guessing the solution takes the form of a pendulum solution and comparing the solutions together to create new information is cool.\\

\section{Problem Four: }
\textbf{The figure below shows a simple pendulum (mass m, length l) whose point of support P is attached to the edge of a wheel (center O, radius R) that is forced to rotate at a fixed angular velocity $\omega$. At t=0, the point P is level with O on the right. Write the Lagrangian and find the equation of motion for the angle $\phi$. [Hint: Be careful writing down the kinetic energy T. A safe way to get the velocity right is to write down the position of the bob at time t, and then differentiate.] Check that your answer makes sense in the special case that $\omega=0$.}
\begin{align*}
  x &= R\cos(\omega t) + l\sin(\phi)\\
  y &= R\sin(\omega t) - l\cos(\phi)\\
  \dot{x} &= -\omega R\sin(\omega t) + l\cos(\phi)\dot{\phi}\\
  \dot{y} &= \omega R\cos(\omega t) + l\sin(\phi)\dot{\phi}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}mv^2\\
  &= \frac{1}{2}m\left(\omega^2R^2\sin^2(\omega t) + l^2\cos^2(\phi)\dot{\phi}^2 - 2\omega Rl\sin(\omega t)\cos(\phi)\dot{\phi} + \omega^2 R^2\cos^2(\omega t) + l^2\sin^2(\phi)\dot{\phi}^2 + 2\omega Rl\cos(\omega t)\sin(\phi)\dot{\phi}\right)\\
  &= \frac{1}{2}m\left(\omega^2R^2 + l^2\dot{\phi}^2 + 2\dot{\phi}\omega Rl\left(\cos(\omega t)\sin(\phi) - \sin(\omega t)\cos(\phi)\right)\right)\\
  &= \frac{1}{2}m\left(\omega^2R^2 + l^2\dot{\phi}^2 + 2\dot{\phi}\omega Rl\sin(\phi - \omega t)\right)\\
  U &= mgy = mg\left(R\sin(\omega t) - l\cos(\phi)\right)\\
  L &= \frac{1}{2}m\left(\omega^2R^2 + l^2\dot{\phi}^2 + 2\dot{\phi}\omega Rl\sin(\phi - \omega t)\right) - mg\left(R\sin(\omega t) - l\cos(\phi)\right)\\
  &= \frac{m\omega^2R^2}{2} + \frac{ml^2\dot{\phi}^2}{2} + m\dot{\phi}\omega Rl\sin(\phi - \omega t) - mg\left(R\sin(\omega t) - l\cos(\phi)\right)\\
  \vspace{1cm}\\
  &\frac{dL}{d\phi} - \frac{d}{dt}\frac{dL}{d\dot{\phi}} = 0\\
  &\frac{dL}{d\phi} = m\dot{\phi}\omega Rl\cos(\phi - \omega t) - mgl\sin(\phi)\\
  &\frac{dL}{d\dot{\phi}} = ml^2\dot{\phi} + m\omega Rl\sin(\phi - \omega t)\\
  &\frac{d}{dt}\frac{dL}{d\dot{\phi}} = ml^2\ddot{\phi} - m\omega Rl\cos(\phi - \omega t)\left(\dot{\phi} - \omega\right)\\
  &m\dot{\phi}\omega Rl\cos(\phi - \omega t) - mgl\sin(\phi) - ml^2\ddot{\phi} + m\omega Rl\cos(\phi - \omega t)\left(\dot{\phi} - \omega\right) = 0\\
  &\omega^2R\cos(\phi - \omega t) + g\sin(\phi) = l\ddot{\phi}\\
\end{align*}

This solution makes sense because, for the case where $\omega = 0$, the equation would be just $g\sin(\phi) = l\ddot{\phi}$, which is the equation for a pendulum. This was a cool problem, I feel like I'm getting experienced at using Lagrangians. Applying the equation of tangential velocity was also something I had to remember how to do. I think it's interesting that coming at the problem from different angles (cartesian coordinates in one direction or another) will yield the same answer.\\

>>>>

\title{Classical Mechanics Assignment 8}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One: }
\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}

\section{Problem Two: }
\textbf{The potential for an non-harmonic oscillator is:}

\begin{equation*}
  U(x) = \frac{kx^2}{2} + \frac{bx^4}{4}\\
\end{equation*}

\textbf{where k and b are positive constants. Find Hamilton’s equations of motion. No final discussion needed.}

\section{Problem Three: }
\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal). Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}

\section{Problem Four: }
\textbf{A thin hoop of radius R and mass M oscillates in its own plane hanging from a single fixed point. Attached to the hoop is a small mass M constrained to move (in a frictionless manner) along the hoop. The inertia of the loop for rotation around a point on the loop is $I=2MR^2$. Consider only small oscillations, and show that the eigenfrequencies are:}

\begin{equation*}
  \omega_1 = \sqrt{\frac{2g}{R}}
\end{equation*}

\begin{equation*}
  \omega_2 = \sqrt{\frac{g}{2R}}
\end{equation*}
